---
title: "written_assignment"
author: "Mohamed Ismael, Mohamed Elgendy"
date: "2022-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
  library(tidyverse)
```

```{r}
  qog_std <- read.csv("dataset/qog_std_cs_jan22.csv")



qog_std

```
```{r}
 qog_std$oecd_c
```
Basic Human Needs (SPI) (spi_bn)
Human Capital Index (egov_hci)
COVID
Political Rights (fh_pr)
Paid Bribe: Police (gcb_bpol)
Total bribery rate, total population (gcb_br)
Accountability and Transparency (iiag_acc)
Anti-corruption (iiag_corr)
Number of previous democratic breakdowns (bmr_dembr)
Number of COVID-19 cases reported (jht_ccc) 436
Number of COVID-19 deaths reported (jht_ccd)
Life Expectancy, Both sexes, Age 1-4 years (ihme_lifexp_0104t)
Total fertility rates (oecd_fertility_t1)
Health (iiag_he)
Road fatalities (oecd_rddeath_t1)
Adult population smoking daily (oecd_smoke_g1)
Alcohol consumption per capita (in litres) (who_alcohol10)
Suicide Rate (per 100,000 population), Total (who_suit)
Employment rate total 15-64 (eu_empy1564t)
Youths not in education or employment 15-19 (oecd_socexclus_t1a)
Youths not in education or employment 20-24 (oecd_socexclus_t1b)
liberal democracy index (vdem_libdem)

# Main question : How has COVID-19 effected education and mental health? (suicide rate, alcohol consumptions)


```{r}
covid_effect <- qog_std %>%
  select(
    jht_ccc,
    jht_ccd,
    who_alcohol10,
    who_suit,
    eu_empy1564t,
    who_let,
    egov_hci,
    vdem_libdem,
    eu_isiuhlt,
    vdem_corr,
    bti_eo,
    fh_pr
  )%>%
  rename(
    covid_deaths = jht_ccc,
    covid_cases = jht_ccd,
    alcohol_consumption = who_alcohol10,
    suicide_rate = who_suit,
    employment_rate = eu_empy1564t,
    life_expectancy = who_let,
    basic_human_needs = egov_hci,
    liberal_democracy = vdem_libdem,
    political_rights = fh_pr,
    seeking_health_info = eu_isiuhlt,
    political_corr = vdem_corr,
    equal_oppor = bti_eo
  )

covid_effect
```
```{r}
  ggplot(data = covid_effect) +
  geom_smooth(mapping = aes(x = liberal_democracy, y = political_corr))
```






```{r}
  ggplot(data = covid_effect) +
  geom_smooth(mapping = aes(x = liberal_democracy, y = equal_oppor))


```




