---
title: "written_assignment"
author: "Mohamed Ismael, Mohamed Elgendy"
date: "2022-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
  library(tidyverse)
```

```{r}
  qog_std <- read.csv("dataset/qog_std_cs_jan22.csv")



qog_std

```
```{r}
  ggplot(data = qog_std) +
  geom_point(mapping = aes(x = spi_bn, y = egov_hci))
```
Basic Human Needs (SPI) (spi_bn)
Human Capital Index (egov_hci)
COVID
Political Rights (fh_pr)
Paid Bribe: Police (gcb_bpol)
Total bribery rate, total population (gcb_br)
Accountability and Transparency (iiag_acc)
Anti-corruption (iiag_corr)
Number of previous democratic breakdowns (bmr_dembr)
Number of COVID-19 cases reported (jht_ccc) 436
Number of COVID-19 deaths reported (jht_ccd)
Life Expectancy, Both sexes, Age 1-4 years (ihme_lifexp_0104t)
Total fertility rates (oecd_fertility_t1)
Health (iiag_he)
Road fatalities (oecd_rddeath_t1)
Adult population smoking daily (oecd_smoke_g1)
Alcohol consumption per capita (in litres) (who_alcohol10)
Suicide Rate (per 100,000 population), Total (who_suit)

